---
title: "More Info - Trends in U.S. Preclearance Notices"
author: "Fulton Jackson, Eion Blanchard"
date: "Last updated on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# Step 0: Setup
# Load relevant libraries
#data analysis
library(xlsx)
library(tidyverse)
library(stringr)
library(plyr)
library(dplyr)
library(readr)
#mapping
library(ggplot2)

# Initialize constants
STATES <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming")
```

### Question
insert project statement (question) here

### Datasets

* Notices of Section 5 activity under the Voting Rights Act of 1965, as amended
  + Link: https://www.justice.gov/crt/archive-notices-section-5-activity-under-voting-rights-act-1965-amended
  + Used data for September 27, 2010 to July 30, 2012
  + Employed .xls version
  + Noticing state, county, and subjurisdiction logged alongside:
    - submission ID
    - action date
    - changes enacted

```{r, eval = FALSE}
# Step 1: Download datasets
dates <- seq(as.Date("2010-09-27"), as.Date("2012-07-30"), by = 1)
urls <- paste("https://www.justice.gov/crt/about/vot/notices/vnote",
              str_sub(dates, 6, 7),
              str_sub(dates, 9, 10),
              str_sub(dates, 3, 4),
              ".xls", sep = "")
oldw <- getOption("warn")
options(warn = -1)
for (url in urls) {
    tryCatch(download.file(url,
                           paste("./data/", str_sub(url, 47), sep = ""),
                           quiet = FALSE),
             error = function(e) print("")) #consider tracking the successful urls   
}
options(warn = oldw)
```

```{r}
# Step 2: Import and clean the data
read_and_tidy <- function(fileName) {
  file_messy <- read.xlsx(fileName, sheetName = "Sheet1")
  file <- file_messy %>%
    select(1, 4, 7, 11, 15, 18, 21, 25) %>%
    unite(NA..23, NA..23, NA..19) %>%
    dplyr::rename(state = NA.,
                  county = NA..3,
                  subjurisdiction = NA..5,
                  submission = NA..9,
                  action_date = NA..13,
                  action = NA..16,
                  changes = NA..23) %>%
    filter(state %in% STATES) %>%
    mutate(action_date = as.Date(action_date, "%m/%d/%Y"),
           changes = str_replace(changes, "_NA", ""),
           changes = str_replace(changes, "NA_", ""))
  
  levels <- levels(file$county)
  levels[length(levels) + 1] <- "None"
  file$county <- factor(file$county, levels = levels)
  file$county[is.na(file$county)] <- "None"
  levels <- levels(file$subjurisdiction)
  levels[length(levels) + 1] <- "None"
  file$subjurisdiction <- factor(file$subjurisdiction, levels = levels)
  file$subjurisdiction[is.na(file$subjurisdiction)] <- "None"
  return(file)
}
```

```{r}
# Step 3: Join datasets into a single object
myfiles <- list.files(path = "./data", pattern = "*.xls", full.names = TRUE)
data <- ldply(myfiles, read_and_tidy)
```

```{r}
# Step 4: Enrich the data
#consolidate the described changes and encode the most common
#account for inconsistencies/generalizations in original dataset
data <- data %>%
  mutate(code = ifelse(str_detect(changes, paste(c("edistricting plan", "onsolidation"),collapse = '|')), 1,
                ifelse(str_detect(changes, paste(c("General election", "Primary election", "Runoff", "Referendum", "Election", "f election", "oint", "election proced"), collapse = '|')), 2,
                ifelse(str_detect(changes, paste(c("Absentee", "oting method", "oter assist"), collapse = '|')), 3,
                ifelse(str_detect(changes, "olling place"), 4,
                ifelse(str_detect(changes, "recinct"), 5,
                ifelse(str_detect(changes, paste(c("Annexation", "annexed area"), collapse = '|')), 6,
                ifelse(str_detect(changes, "egistration"), 7,
                8 ))))))))
                # ifelse(str_detect(changes, "dministration"), 9,
                # ifelse(str_detect(changes, "Term of office"), 10,
                # ifelse(str_detect(changes, "Abolishment"), 11,
                # ifelse(str_detect(changes, "Ballot format"), 12,
                # ifelse(str_detect(changes, "Boundary changes"), 13,
                # ifelse(str_detect(changes, "Campaign financing"), 14,
                # ifelse(str_detect(changes, "Candidate qualification"), 15,
                # ifelse(str_detect(changes, "Compensation"), 16,
                # ifelse(str_detect(changes, "Concurrent terms"), 17,
                # ifelse(str_detect(changes, "Creation"), 18,
                # ifelse(str_detect(changes, "Deannexation"), 19,
                # ifelse(str_detect(changes, "Dissolution"), 20,
                # ifelse(str_detect(changes, "Districting"), 21,
                # ifelse(str_detect(changes, "Establishment"), 22,
                # ifelse(str_detect(changes, "Form"), 23,
                # ifelse(str_detect(changes, "Implementation Schedule"), 24,
                # ifelse(str_detect(changes, "Incorporation"), 25,
                # ifelse(str_detect(changes, "Initiative, Ref"), 26,
                # ifelse(str_detect(changes, "Limited"), 27,
                # ifelse(str_detect(changes, "Majority"), 28,
                # ifelse(str_detect(changes, "Nonpartison"), 29,
                # ifelse(str_detect(changes, "Number of officials"), 30,
                # ifelse(str_detect(changes, "Numbered positions"), 31,
                # ifelse(str_detect(changes, "Powers and duties"), 32,
                # ifelse(str_detect(changes, "Redistricting procedures"), 33,
                # ifelse(str_detect(changes, "Staggered terms"), 34,
                # ifelse(str_detect(changes, "Transfer of powers"), 35,
                # ifelse(str_detect(changes, "Voting qualification"), 36,
                # ifelse(str_detect(changes, "Nominating"), 37,
                # ifelse(str_detect(changes, "Political activity"), 38,
                # ifelse(str_detect(changes, "Procedures"), 39,
                # ifelse(str_detect(changes, "Bilingual"), 40,
                # ifelse(str_detect(changes, "selection"), 41,
                # ifelse(str_detect(changes, "staging"), 42,
                # ifelse(str_detect(changes, "Plurality"), 43,
                # ifelse(str_detect(changes, "Purge"), 44, 45)
                # ))))))))))))))))))))))))))))))))))))))))))))
```

### Graphics
Below, we plot the preclearance notice changes and their trends in frequency.

```{r}
# Step 5: Plot mosaics of change notices
#employ only ggplot2, not ggmosaic
#prepare the data for national plotting
data_p <- data %>%
  mutate(date = format(action_date, "%y %m")) %>%
  group_by(date, code) %>%
  dplyr::summarize(s = sum(code)) %>%
  mutate(n = s / code) %>% # assumes that code values are nonzero
  select(-s)
widths <- data_p %>%
  group_by(date) %>%
  dplyr::summarize(xwidth = sum(n))
data_p <- data_p %>%
  left_join(widths, by = "date") %>%
  mutate(month = format(as.Date(paste(date, "01", sep = " "), "%Y %m %d"), "%b '%y"))
data_p$month = factor(data_p$month,
                      levels=c("Sep '10", "Oct '10", "Nov '10", "Dec '10",
                               "Jan '11", "Feb '11", "Mar '11", "Apr '11", "May '11", "Jun '11",
                               "Jul '11", "Aug '11", "Sep '11", "Oct '11", "Nov '11", "Dec '11",
                               "Jan '12", "Feb '12", "Mar '12", "Apr '12", "May '12", "Jun '12", "Jul '12"))
#plot the mosaic for combined national counts
data_p %>%
  print()%>%
  ggplot(aes(x = factor(month), y = n)) +
  geom_col(aes(width = xwidth,
               fill = factor(code)),
           colour = "white",
           size = .5,
           position = position_fill(reverse = TRUE)) +
  facet_grid(~ month, space = "free", scales = "free", switch = "x") +
  scale_x_discrete(name = "Month submitted") +
  scale_y_continuous(labels = scales::percent) +
  theme(strip.text.x = element_text(size = 6.5, angle = 0),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing = unit(0, "pt")) +
  scale_fill_discrete(name = "",
                      breaks = c("1", "2", "3", "4", "5", "6", "7", "8"),
                      labels = c("redistricting plan", "election procedures", "voting method", "polling place", "precinct", "annexation", "voter registration", "other"),
                      guide = guide_legend(reverse=TRUE)) +
  labs(title = "Preclearance Submissions for the Voting Rights Act of 1965, Section 5")
```

```{r}
#prepare the data for statewide plotting
data_ps <- data %>%
  mutate(date = format(action_date, "%y %m")) %>%
  group_by(date, code, state) %>%
  dplyr::summarize(s = sum(code)) %>%
  mutate(n = s / code) %>% # assumes that code values are nonzero
  select(-s)
widths <- data_ps %>%
  group_by(date, state) %>%
  dplyr::summarize(xwidth = 1)
data_ps <- data_ps %>%
  left_join(widths, by = c("date", "state")) %>%
  mutate(month = format(as.Date(paste(date, "01", sep = " "), "%Y %m %d"), "%b '%y"))
data_ps$month = factor(data_ps$month,
                      levels=c("Sep '10", "Oct '10", "Nov '10", "Dec '10",
                               "Jan '11", "Feb '11", "Mar '11", "Apr '11", "May '11", "Jun '11",
                               "Jul '11", "Aug '11", "Sep '11", "Oct '11", "Nov '11", "Dec '11",
                               "Jan '12", "Feb '12", "Mar '12", "Apr '12", "May '12", "Jun '12",
                               "Jul '12"))
#plot the mosaic for combined state counts
data_ps %>%
  print()%>%
  ggplot(aes(x = factor(month), y = n)) +
  geom_col(aes(width = xwidth,
               fill = factor(code)),
           colour = "white",
           size = .5,
           position = position_fill(reverse = TRUE)) +
  facet_grid(~ month, space = "free", scales = "free", switch = "x") +
  scale_x_discrete(name = "Month submitted") +
  scale_y_continuous(labels = scales::percent) +
  theme(strip.text.x = element_text(size = 6.5, angle = 0),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing = unit(0, "pt")) +
  scale_fill_discrete(name = "",
                      breaks = c("1", "2", "3", "4", "5", "6", "7", "8"),
                      labels = c("redistricting plan", "election procedures", "voting method", "polling place", "precinct", "annexation", "voter registration", "other"),
                      guide = guide_legend(reverse=TRUE)) +
  labs(title = "Preclearance Submissions for the Voting Rights Act of 1965, Section 5 (by state)") +
  facet_wrap(~state)
```

```{r}
# Step 6: Modelling trends in the preclearance change notices
#prepare data for regression
data_t <- data_p %>%
  select(-xwidth) %>%
  spread(key = code, value = n) %>%
  dplyr::rename(redistricting = 3, election = 4, method = 5, polling = 6, precinct = 7, annex = 8, registration = 9, other = 10)
data_t[is.na(data_t)] <- 0
```

```{r}
#model redistricting notices in terms of precinct notices
model <- glm(formula = redistricting ~ precinct, data = data_t)
summary(model)

# plot(redistricting ~ precinct, data = data_t)
# abline(model)

data_t %>%
  ggplot(aes(y = 'redistricting', x = precinct)) +
  geom_point(alpha = 0.3) +
  labs(y = "Count of redistricting notices", x = "Count of precinct notices") +
  geom_smooth(method = "lm", se = FALSE)
```

A solid relationship exists between the count of redistricting notices and the count of precinct change notices, by month. This makes intuitive sense, since redistricting plans often included modified precincts in order to satisfy the "one person, one vote" population stipulation.

```{r}
#model and plot redistricting notices in terms of other relevant counts
model <- glm(formula = redistricting ~ polling + method, data = data_t)
summary(model)
plot(model)
```

Without the predictor of precinct change notices, we may gain an accurate estimate for the count of redistricting notices from the polling place and the voting method notices.

```{r}
#model and plot redistricting notices in terms of election procedure notcies
model <- glm(formula = redistricting ~ election, data = data_t)
summary(model)
plot(model)
```

A weak correlation exists between the counts of redistricting and election procedure change notices.
